\documentclass[12pt]{article}

\newcommand{\isdir}{\ensuremath{\mathit{isdir}}}
\newcommand{\dir}{\ensuremath{\mathrm{dir}}}
\newcommand{\failed}{\ensuremath{\mathit{failed}}}
\newcommand{\To}[1]{\stackrel{#1}{\longrightarrow}}
\newcommand{\agreeAt}[1]{\sim_{#1}}
\newcommand{\agreeAround}[1]{\sim_{\langle #1\rangle}}
\newcommand{\AgreeAt}[1]{\approx_{#1}}
\newcommand{\AgreeAround}[1]{\approx_{\langle #1\rangle}}
\newcommand{\convergesTo}[2]{\stackrel{#1}{\Longrightarrow}_{#2}}

\begin{document}

\title{Synchronizer specification}
\author{Trevor Jim}
\maketitle


\section{Preliminaries}

\begin{itemize}
\item Define $A \agreeAt{p} B$ ($A$ agrees with $B$ at $p$) iff
  $\isdir_{AB}(p)$ or $A(p)=B(p)$.
\item Define $A \agreeAround{p} B$ ($A$ agrees with $B$ around $p$)
  iff $A \agreeAt{q} B$ for every prefix $q\leq p$
  and every extension $q\geq p$.
\item States are elements of an abstract type with predicate
  $\AgreeAt{p}$ between states and filesystems.  We use $M$ to range
  over states.  Say that $M$ is forgetful at $p$ iff
  $M\not\AgreeAt{p}A$ for all filesystems $A$.  We require states to
  satisfy the following properties:
  \begin{enumerate}
  \item If $M\AgreeAt{p}A$ and $M\AgreeAt{p}B$, then $A\agreeAt{p}B$.
  \item If $M$ is forgetful at $p$, then $M$ is forgetful at $q$ for
    all paths $q\geq p$.
  \item For every $M$ there exists a filesystem $A$ such that for all $p$,
    either $M\AgreeAt{p}A$ or $M$ is forgetful at $p$.
  \end{enumerate}
  I believe that only (1) is strictly necessary for the specification,
  and properties (2) and (3) enable an efficient algorithm.

  We can extend $\AgreeAt{p}$ to relate states and states as follows:
  \begin{itemize}
  \item $M\AgreeAt{p}M'$
    iff for all filesystems $A$, $M\AgreeAt{p}A$ iff $M'\AgreeAt{p}A$
  \end{itemize}
  The predicate $\AgreeAround{p}$ is defined between states and
  filesystems, and states and states, in the obvious way.
\end{itemize}

Note that we can have $M\AgreeAt{p}A$, $A\agreeAt{p}B$,
and $M\not\AgreeAt{p}B$ all at the same time.

\section{The first specification}

\begin{verbatim}
a synchronizer S comprises
    a synchronization relation
       snc : FS * FS * STATE * FS * FS * STATE * BOOL

a synchronizer S is CORRECT if
  forall A,B,A',B' : FS
  forall M,M' : STATE
  forall failed : BOOL
  if snc(A,B,M, A',B',M', failed), then
    for all p.
\end{verbatim}

\begin{enumerate}
\item If the synchronizer changes the state, it does so because the
  replicas have been synchronized:
  \begin{quote}
    if $M'\not\AgreeAt{p} M$,
    then $M'\AgreeAt{p} A'$, $M'\AgreeAt{p} B'$, and $A'\agreeAt{p} B'$
  \end{quote}
  
\item If the synchronizer changes one replica, it does so to bring it
  in sync with a change in the other:
  \begin{quote}
    if $A'\not\agreeAt{p} A$,
    then $A'\agreeAt{p} B'\agreeAt{p} B\not\AgreeAt{p} M$

    if $B'\not\agreeAt{p} B$,
    then $B'\agreeAt{p} A'\agreeAt{p} A\not\AgreeAt{p} M$
  \end{quote}
      
\item If the user makes a change the synchronizer leaves it alone:
  \begin{quote}
    if $A\not\AgreeAt{p} M$,
    then $A'\agreeAt{p} A$

    if $B\not\AgreeAt{p} M$,
    then $B'\agreeAt{p} B$
  \end{quote}
  
\item If only one side changed, then the change should be propagated;
  if not, we screwed up:
  \begin{quote}
    if ($A\AgreeAround{p} M$ or $B\AgreeAround{p} M$),
    then ($A'\agreeAt{p} B'$ or $\failed$)
  \end{quote}
  
\item If the replicas are synchronized but the synchronizer doesn't
  realize it, we screwed up:
  \begin{quote}
    if $A'\agreeAt{p} B'$,
    then
    ($M'\AgreeAt{p} A'$ and $M'\AgreeAt{p} B'$) or $\failed$
  \end{quote}

\end{enumerate}


\subsection{Comments}

There are a bunch of other conditions that should hold; I think these
are all implied by the above.

\begin{enumerate}
\item If the user didn't change $A$, the synchronizer leaves $B$
  alone, and propagates any change in $B$ to $A$ or fails:
  \begin{quote}
    if $A\AgreeAround{p} M$,
    then $B'\agreeAt{p} B$ and
    ($A'\agreeAt{p} B'$ or $\failed$)
  \end{quote}
  
\item If the replicas are already in sync, the synchronizer leaves
  them alone:
  \begin{quote}
    if $A\agreeAt{p} B$,
    then $A'\agreeAt{p} B'$
  \end{quote}
  
\item If the replicas end up in sync, the synchronizer has left them
  in a state determined by the user:
  \begin{quote}
    if $A'\agreeAt{p} B'$,
    then $A'\agreeAt{p} A$ or $A'\agreeAt{p} B$
  \end{quote}
  
\item If the replicas don't end up in sync, the synchronizer has made
  no change:
  \begin{quote}
    if $A'\not\agreeAt{p} B'$,
    then $A'\agreeAt{p} A$ and $B'\agreeAt{p} B$
  \end{quote}
  
\item If a conflict has been introduced the synchronizer does nothing:
  \begin{quote}
    if $A\not\AgreeAt{p} M$ and $B\not\AgreeAt{p} M$ and $A\not\agreeAt{p} B$,
    then $A'\agreeAt{p} A$ and $B'\agreeAt{p} B$ and $M'\AgreeAt{p} M$
  \end{quote}
  
\item The synchronizer touches at most one side:
  \begin{quote}
    $A'\agreeAt{p} A$ or $B'\agreeAt{p} B$
  \end{quote}
  
\item The possible values for $A'$ and $B'$ come from $A$ and $B$ only
  \begin{quote}
    ($A'\agreeAt{p} A$ or $A'\agreeAt{p} B$) and ($B'\agreeAt{p} A$ or
    $B'\agreeAt{p} B$)
  \end{quote}
\end{enumerate}

We need to prevent cases like

\begin{verbatim}
    A - 1  A' - 2    A - 1  A' - 1
    B - 2  B' - 1    B - 2  B' - 1
    M - 3  M' - 3    M - 3  M' - 1

    A - 1  A' - 3    A - 1  A' - 1
    B - 2  B' - 3    B - 2  B' - 1
    M - 3  M' - 3    M - 1  M' - 1
\end{verbatim}


There are also some things that do not hold; these are from earlier
failed attempts at giving a simpler specification.

\begin{enumerate}
\item 
  \begin{quote}
    if ($A\AgreeAt{p} M$ or $B\AgreeAt{p} M$),
    then ($A'\agreeAt{p} B'$ or $\failed$)
  \end{quote}
  The counterexample due to Matthieu is
  \begin{quote}
    $M: \dir{}\To{x}\dir{}\To{y}\textrm{file}_1$
    
    $A: \dir{}\To{x}\dir{}\To{y}\textrm{file}_2$
    
    $B: \dir{}\To{x}\textrm{file}_3$
  \end{quote}
  Here $A\AgreeAt{x}M$, so it looks like the change at $B$ should be
  propagated but in fact there is a conflict.

\item The same counterexample show that the following also does not hold.
  \begin{quote}
    if $\forall q\leq p.\ A\AgreeAt{q} M$,
    then ($A'\agreeAt{p} B'$ or $\failed$)

    if $\forall q\leq p.\ B\AgreeAt{q} M$,
    then ($A'\agreeAt{p} B'$ or $\failed$)
  \end{quote}

\item Similarly, the following does not hold:
  \begin{quote}
    if $\forall q\geq p.\ A\AgreeAt{q} M$,
    then ($A'\agreeAt{p} B'$ or $\failed$)

    if $\forall q\geq p.\ B\AgreeAt{q} M$,
    then ($A'\agreeAt{p} B'$ or $\failed$)
  \end{quote}
  The counterexample is
  \begin{quote}
    $M: \dir{}\To{x}\textrm{file}_1$
    
    $A: \dir{}\To{x}\textrm{file}_2$
    
    $B: \dir{}\To{x}\dir{}\To{y}\textrm{file}_3$
  \end{quote}
  Here $A\AgreeAt{q}M$ for all $q\geq x.y$, so it looks like the change
  at $B$ should be propagated but in fact there is a conflict.
\end{enumerate}


\section{Multiple replicas}

\begin{itemize}
\item 
Use $R,S,T$ to range over lists of filesystems.

\item We extend $\agreeAt{p}$ to lists of file systems:
  \begin{quote}
    $R \agreeAt{p} S$ iff $\forall A\in R,B\in S$ we have $A \agreeAt{p} B$
  \end{quote}
  Note that $\agreeAt{p}$ is not reflexive on lists of filesystems.
  The relation $\agreeAround{p}$ is extended
  similarly, as are the relations involving states.
  
\item We write $M \convergesTo{A}{p} R$ iff the following three
  conditions hold:
  \begin{enumerate}
  \item $A\in R$;
  \item $M\not\AgreeAt{p} A$; and
  \item $\forall B\in R$, either $B\AgreeAround{p} M$ or
    $B\agreeAround{p} A$.
  \end{enumerate}
\end{itemize}

\begin{verbatim}
A synchronizer Sync is CORRECT if
  forall R={A1 .. An} : FS list,
         R'={A1' .. An'} : FS list,
         M,M' : STATE
  if Sync(R,R',M,M',failed), then for all p:
\end{verbatim}
\begin{enumerate}
\item If the synchronizer changes the state, it does so because the
  replicas have been synchronized:
  \begin{quote}
    if $M'\not\AgreeAt{p} M$,
    then $M'\AgreeAt{p} R'$
  \end{quote}
  
\item If the synchronizer changes a replica, it does so to bring it in
  sync with a change in another replica:
  \begin{quote}
    $\forall i\leq n$,
    if $A_i'\not\agreeAt{p} A_i$,
    then for some $A$,  $M\convergesTo{A}{p} R$  and $A_i'\agreeAt{p} A$
  \end{quote}
  
\item If the replicas can be synchronized at $p$, the synchronizer
  should make that happen:
  \begin{quote}
    if $M\convergesTo{A}{p} R$ for some $A$,
    then ($R'\agreeAt{p} R'$ or $\failed$)
  \end{quote}
  
\item If the replicas end up synchronized, the synchronizer should
  realize this:
  \begin{quote}
    if $R'\agreeAt{p} R'$, then $ M'\AgreeAt{p}R'$ or $\failed$
  \end{quote}
\end{enumerate}


\end{document}
